HBCC - ACC - STG - FEATURECOUNTS - GRCh38 - CQN - Sparse Model

### 2020-01-05 cmc-rnaseq-de/HBCC_ACC-STG_FEATURECOUNTS_GRCh38_CQN_Sparse.Rmd
fix contrasts

plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, "SampleID")) %>%
  dplyr::mutate(label = SampleID) %>% 
  dplyr::mutate(label = ifelse((label %in% outliers), label, NA))

p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(shape=Dx.Tissue, size=ageOfDeath))
p <- p + theme_bw() %+replace% theme(legend.position="right")
p <- p + geom_text(aes(label= label), size=4, hjust=0)
p
# Plot abberent distribution of logcpm counts
tmp1 = voom(PROCESSED_COUNTS$filteredExprMatrix$counts)$E %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  tidyr::gather(SampleID, logCPM, -ensembl_gene_id) %>%
  left_join(COVARIATES %>%
              rownameToFirstColumn('SampleID'))
# p = ggplot(tmp1 %>%
#              dplyr::filter(SampleID %in% indToRemove),
#            aes(x = logCPM, color = SampleID)) + geom_density() 
# p = p + theme_bw() %+replace% theme(legend.position = 'none') + facet_grid(Dx~., scale = 'free')
# p

@@

p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(shape=Dx.Tissue, size=ageOfDeath))
p <- p + theme_bw() %+replace% theme(legend.position="top", plot.title = element_text(hjust=0.5, size = 15)) + ggtitle ("PCA MSSM/Penn/Pitt Cohort") + 
  guides(colour = guide_legend(nrow = 2),
         shape = guide_legend(nrow = 2),
         size = guide_legend(nrow = 2))
p
# p <- p + geom_text(aes(label= SampleID), size=4, hjust=0)

@@

p <- ggplot(plotdata, aes(MDS1,MDS2))
p <- p + geom_point(aes(shape=Dx.Tissue, size=ageOfDeath))
p <- p + theme_bw() %+replace% theme(legend.position="top", plot.title = element_text(hjust=0.5, size = 15)) + ggtitle ("MDS MSSM/Penn/Pitt Cohort") + 
  guides(colour = guide_legend(nrow = 2),
         shape = guide_legend(nrow = 2),
         size = guide_legend(nrow = 2))
p
```
Tree based clustering of samples
```{r decompse.normalise.data.1, fig.height=6, fig.width=10, results='asis'}
# Eucledian tree based analysis
COVARIATES.tmp = data.matrix(COVARIATES[,c("Reported_Gender", "Dx.Tissue")])
COVARIATES.tmp[is.na(COVARIATES.tmp)] = 0

@@

>>>>

plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, "SampleID")) %>%
    tidyr::separate(Dx.Tissue, c('Dx','Tissue'), sep = '_') %>%
  dplyr::mutate(label = SampleID) %>% 
  dplyr::mutate(label = ifelse((label %in% outliers), label, NA))

p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(shape=Tissue, size=ageOfDeath))
p <- p + theme_bw() + theme(legend.position="right") + facet_grid(Tissue~.)
p <- p + geom_text(aes(label= label), size=4, hjust=0)
p

# Plot abberent distribution of logcpm counts
tmp1 = voom(PROCESSED_COUNTS$filteredExprMatrix$counts)$E %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  tidyr::gather(SampleID, logCPM, -ensembl_gene_id) %>%
  left_join(COVARIATES %>%
              rownameToFirstColumn('SampleID') %>%
              tidyr::separate(Dx.Tissue, c('Dx', 'Tissue'), sep = '_'))

p = ggplot(tmp1 %>%
             dplyr::filter(SampleID %in% outliers),
           aes(x = logCPM, color = SampleID)) + geom_density() 
p = p + theme(legend.position = 'top') + facet_grid(Dx+.~Tissue, scale = 'free')
p

@@

p <- ggplot(plotdata, aes(x=PC1, y=PC2))
p <- p + geom_point(aes(shape=Dx.Tissue, size=ageOfDeath))
p <- p + theme_bw() %+replace% theme(legend.position="top", plot.title = element_text(hjust=0.5, size = 15)) + ggtitle ("PCA HBCC ACC-STG Cohort") + 
  guides(colour = guide_legend(nrow = 2),
         shape = guide_legend(nrow = 2),
         size = guide_legend(nrow = 2))
p
# p <- p + geom_text(aes(label= SampleID), size=4, hjust=0)

@@

p <- ggplot(plotdata, aes(MDS1,MDS2))
p <- p + geom_point(aes(shape=Dx.Tissue, size=ageOfDeath))
p <- p + theme_bw() %+replace% theme(legend.position="top", plot.title = element_text(hjust=0.5, size = 15)) + ggtitle ("MDS HBCC ACC-STG Cohort") + 
  guides(colour = guide_legend(nrow = 2),
         shape = guide_legend(nrow = 2),
         size = guide_legend(nrow = 2))
p
```
Tree based clustering of samples
```{r decompse.normalise.data.1, fig.height=6, fig.width=10, results='asis'}
# Eucledian tree based analysis
COVARIATES.tmp = data.matrix(COVARIATES[,c("Reported_Gender", "Dx.Tissue")])
COVARIATES.tmp[is.na(COVARIATES.tmp)] = 0

@@

plotdata <- left_join(plotdata, rownameToFirstColumn(COVARIATES, 'SampleID')) %>%
  tidyr::separate(Dx.Tissue, c('Dx','Tissue'), sep = '_')
